---
title: "p8105_hw2_qw2331"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

library(readxl)
library(Hmisc)
```

## Problem 1

**Read & clean the Mr. Trash Wheel sheet**
```{r}
trash_df <- 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N534") %>% 
  janitor::clean_names() %>% 
  filter(is.na(dumpster) == FALSE) %>% 
  mutate(sports_balls = round(sports_balls))
```

**Read & clean precipitation data for 2018 and 2019**
```{r}
# Read and clean precipitation data for 2018
precip18_df <- 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018)

# Read and clean precipitation data for 2019
precip19_df <- 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2019)

# Combine two data frame & convert `month`
precip_df <- 
  bind_rows(precip18_df, precip19_df) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month, total)
```

**A brief description about the data**  
There are ``r nrow(trash_df)`` observations in the Mr. Trash Wheel dataframe after cleanness and ``r nrow(precip_df)`` in the second one after combination. Among all ``r ncol(trash_df)`` variables in dataset `trash_df` (``r names(trash_df)``) and ``r ncol(precip_df)`` variables in the dataset `precip_df` (``r names(precip_df)``). The key variables are `month` and `year` since they are in both datasets and can be used to join these two datasets. 

For available data, the total precipitation in 2018 is ``r precip18_df %>% pull(total) %>% sum()`` and the median number of sports balls in a dumpster in 2019 is ``r trash_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()``.


## Problem 2

**Clean `pols-month.csv`**
```{r, message = FALSE}
pols_df <- 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month), 
         day = as.integer(day),
         month = month.name[month],
         president = ifelse(prez_gop == 0, "dem", "gop")
         ) %>% 
  select(-prez_dem & -prez_gop & -day) %>% 
  relocate(year, month, president)
```

**Clean `snp.csv`**
```{r, message = FALSE}
snp_df <- 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         month = month.name[month]) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
```

**Tidy `unemployment.csv` up**
```{r, message = FALSE}
unemploy_df <- 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemploy_pect"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(year = as.integer(year),
         month = match(month, month.abb),
         month = month.name[month])
```

**Join the datasets by merging `snp`, `pols`, and `unemployment`**
```{r}
PolEco_df <- 
  left_join(pols_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemploy_df, by = c("year", "month"))
```

**A brief description about the data**  
The `pols_df` contains `822` observations of `9` variables related to the number of national politicians at any given time;  
The `snp_df` contains `787` observations of `3` variables related to Standard & Poor's stock market index;  
The `unemploy_df` contains `816` observations of `3` variables related to the percentage of unemployment in ant given time.

After merging all the datasets, the resulting dataframe `PolEco_df` contains ``r nrow(PolEco_df)`` obervations of ``r ncol(PolEco_df)`` variables. The variable `year` ranges from ``r min(pull(PolEco_df, year))`` to ``r max(pull(PolEco_df, year))``. And the names of key variables are `year` and `month`.


## Problem 3

**Load `Popular_Baby_Names.csv` and tidy up**
```{r, message = FALSE}
popbaby_df <- 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename(first_name = childs_first_name) %>% 
  mutate(
    gender = tolower(gender),
    ethnicity = tolower(ethnicity),
    first_name = tolower(first_name),
    first_name = capitalize(first_name),
    ethnicity = recode(ethnicity, 
                       "asian and pacific islander" = "asian and paci",
                       "black non hispanic" = "black non hisp",
                       "white non hispanic" = "white non hisp")
  ) %>% 
  distinct()
```


**Produce table showing the rank of female name "Olivia"**
```{r}
olivia_df <- 
  filter(popbaby_df, first_name == "Olivia" & gender == "female") %>% 
  select(-count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  relocate(c(1:3), c(9:4))

olivia_df %>% 
  knitr::kable()
```


**A similar table showing the most popular male name**
```{r}
popmale_df <- 
  filter(popbaby_df, gender == "male" & rank == 1) %>% 
  select(-count & -rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "first_name"
  ) %>% 
  relocate(1, 2, c(8:3))

popmale_df %>% 
  knitr::kable()
```


**Produce a scatter plot**
```{r}
# Select the data for plotting
wnhmale_df <- 
  filter(
    popbaby_df, 
    gender == "male" & ethnicity == "white non hisp" & year_of_birth == 2016)

# Let's plot
wnhmale_df %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Scatterplot",
    subtitle = "Rank VS. Number of childern",
    x = "Rank of a name",
    y = "Number of children with a name"
  )
```

